---
title: "Week 3 Peer-graded Assignment"
author: "Vincent Blain"
date: "3/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Developing Data Products - Week 3 Peer-graded Assignment

Create a web page presentation using R Markdown that features a plot created with Plotly. Host your webpage on either GitHub Pages, RPubs, or NeoCities. Your webpage must contain the date that you created the document, and it must contain a plot created with Plotly. We would love to see you show off your creativity! 

```{r}
library(plotly)
```

### 3D Surface using Plotly and built in mtcars dataset

```{r}
# mtcars matrix
data <- as.matrix(mtcars)

plot_ly(x=colnames(data), y=rownames(data), z = data, type = "surface") %>% layout(margin = list(l=120))
```

### 3D Surface using Plotly and built in mtcars dataset

Add Package "Tigris" to be able to get US State/County FIPS

```{r}
library(tigris)
library(dplyr)
```

Load midwest dataset
```{r}
mh <- midwest
```

Load FIPS data, remove the string " County" from the county column

```{r}
fips <- fips_codes
fips <- data.frame(lapply(fips, function(x) { gsub(" COUNTY", "", toupper(x)) }))
```

Using DPLYR join the Midwest Dataset to Fips to create a dataset that contains the full FIPS code to make easier to generate a map based on FIPS. Use Stringr to remove all white space from the FullFips column to prevent errors.

```{r}
library(dplyr)
library(stringr)

mhfips <- mh %>% inner_join(fips, by = c("state" = "state", "county" = "county"))
mhfips$fullfips <- as.character(str_replace_all(paste(mhfips$state_code, mhfips$county_code), regex("\\s*"), ""))
```

```{r}
library(rjson)

url <- 'https://raw.githubusercontent.com/plotly/datasets/master/geojson-counties-fips.json'
counties <- rjson::fromJSON(file=url)

mhfips$hover <- with(mhfips, paste(state, '<br>', 
                           county, '<br>', 
                           "white", popwhite, 
                           "black", popblack
                           ))

# give state boundaries a white border
l <- list(color = toRGB("white"), width = 2)

# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

#fig <- plot_geo(mhfips, locationmode = 'USA-states')
fig <- plot_ly()
fig <- fig %>% add_trace(
    type="choropleth",
    z = mhfips$poptotal, 
    geojson=counties,
    text = mhfips$hover, 
    locations = mhfips$fullfips,
    colorscale="Viridis",
    marker=list(line=list(
      width=0)
    )
  )
fig <- fig %>% colorbar(title = "Population per county")
fig <- fig %>% layout(
    title = 'Midwest US State Dataset<br>(Hover for breakdown)',
    geo = g
  )

fig
```
